Information within words in English, French and German speech

### 2019-07-04 surprisal/phoneme_writeup.Rmd
all data has counts
histograms

```{r read-data}
all_data <- read_feather(here("Data/all_phonemes.feather"))
```

@@

Through a series of exploratory analyses we tackle this problem in three ways:
1. Computing the surprisal of characters in English words from baserates of character production (unigram) and then conditioning on the prior charaters in each word (bigram and trigram). This analyses roughly captures the difficulty that a listener might have in predicting each character.
2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations of spoken English. 
3. Computing hold-out entropy for individual phonemes in English words. This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position $X$". 
4. Comparing our English by-letter analysis to results for French and German which have different expected information profiles. We expect that French words will have less information at the end of words compared to English, while German words will have more compared to English.

@@

We use the Providence corpus from CHILDES as data for English. In the plots below, each plot represents the information in each letter, divided by words of different lengths. Word lengths are given by number of letters--we show the length above each plot.

@@

```{r plot-helper}
plot_curves <- function(select_language = "English", select_rep = "letter", 
                        select_gram = "Unigram") {
  all_data %>%
    filter(language == select_language, rep == select_rep,
           gram == select_gram) %>%
    filter(word_length %in% 2:6) %>%
    ggplot(aes(x = order, y = mean_s)) +
    facet_grid(. ~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x, " \n n = ", 
                                                         all_data %>% 
                                                           filter(language == select_language,
                                                                  rep == select_rep, 
                                                                  length == x) %>%
                                                           pull(length) %>%
                                                           unique()))) +
    geom_point() +
    geom_line() +
    xlab(paste0(str_to_sentence(select_rep), " Position")) +
    ylab("Average information") + 
    ggtitle(paste0("Information in ", select_language, " ", 
                   select_rep, "s by ",select_rep, " position - ", select_gram))
}
```

@@

@@

@@

>>>>

```{r read-data}
all_data <- read_feather(here("Data/all_data.feather"))
```

@@

Through a series of exploratory analyses we tackle this problem in three ways:

1. Computing the surprisal of characters in English words from baserates of character production (unigram) and then conditioning on the prior charaters in each word (bigram and trigram). This analyses roughly captures the difficulty that a listener might have in predicting each character.
2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations of spoken English. 
3. Computing hold-out entropy for individual phonemes in English words. This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position $X$". 
4. Comparing our English by-letter analysis to results for French and German which have different expected information profiles. We expect that French words will have less information at the end of words compared to English, while German words will have more compared to English.

@@

```{r plot-hist-helper}
plot_hists <- function(select_language = "English", select_rep = "letter") {

    all_data %>%
    filter(language == select_language, rep == select_rep) %>%
    distinct(word_length, n, rep) %>%
    ggplot(aes(x = word_length, y = n)) +
    geom_col(fill = "white", color = "black") +
    xlab(paste0(str_to_sentence(select_rep), " Length")) +
    ylab("Frequency")
}
```

We use the Providence corpus from CHILDES as data for English. In the plots below, each plot represents the information in each letter, divided by words of different lengths. Word lengths are given by number of letters.

```{r plot-english-letter-hist}
plot_hists()
```

@@

```{r plot-curve-helper}
plot_curves <- function(select_language = "English", select_rep = "letter", 
                        select_gram = "Unigram") {
  all_data %>%
    filter(language == select_language, rep == select_rep,
           gram == select_gram) %>%
    filter(word_length %in% 2:6) %>%
    ggplot(aes(x = order, y = mean_s)) +
    facet_grid(. ~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x))) +
    geom_point() +
    geom_line() +
    xlab(paste0(str_to_sentence(select_rep), " Position")) +
    ylab("Average information") + 
    ggtitle(paste0("Information in ", select_language, " ", 
                   select_rep, "s by ",select_rep, " position - ", select_gram))
}
```

@@

```{r plot-english-phone-hist}
plot_hists(select_rep = "phone")
```

@@

```{r plot-french-hist}
plot_hists(select_language = "French")
```

@@

```{r plot-german-hist}
plot_hists(select_language = "German")
```
